Formalizing Stack Safety as a Security Property

### 2021-09-30 submission4/main.tex
re-added section VI; adjusted contributions

Our threat model is very strong, allowing the attacker (i.e., the
callee) to execute arbitrary code, including attempting to smash
the stack to disrupt the program’s control flow. Our stack safety
properties demand that, even in the presence of such attackers, confidentiality
and integrity still apply to data in suspended stack frames
until the callee returns.

We can also, optionally, include a notion of {\em well-bracketed control
flow (WBCF)} as in Skorstengaard et al. \cite{SkorstengaardSTKJFP}---a global requirement that
callees must always return to their immediate caller, if they return
at all.

@@

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security, combine these with an existing notion of well-bracketed
control flow to formalize a {\em stepwise stack-safety}
property, and which captures the intuitive concept of
strict stack safety enforced over each step of execution~(\cref{sec:prelim,sec:lse}).
\item We use property-based random testing to check that this stepwise
property is  enforced by the {\em Depth Isolation} policy
of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}; in the process, we extend the generation by
execution techniques of Hri\c{t}cu et al.\cite{TestingNI:ICFP} to handle nested
properties (\cref{sec:enforcement}).
\item We formalize weaker but more extensional \emph{observational} variants of
  confidentiality and integrity. Both are parameterized over a notion of external
  observation, and are violated only if accessing secrets or overwriting
  data causes a visible change in the system behavior (\cref{sec:lazy}).
\item We show that our definitions can be
extended to describe argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
\end{itemize}
\Cref{sec:relwork,sec:future} discuss related and future work.

@@

Our model is flexible enough to support many realistic uses of the stack.
We intentionally say little or nothing about other structures
such as the heap.  (Heap safety has already been studied, in isolation,
by Azevedo de Amorim et al., and we expect that their correctness condition and ours can be
combined cleanly to characterize correct protection of stack + heap.)

We do not currently support:
1) address-taken locals, as a satisfactory treatment of these seems to require
integration with a more general model of memory safety,
2) exceptions, \sna{I could just write up how to do these...}
3) tail calls,
4) dynamic code generation, which is difficult to integrate with code annotations in general, and
5) concurrency --- we can handle a limited form of coroutines, but
  not full-blown preemptive concurrency.

@@

\medskip
\paragraph*{Case Study: Micro-Policies}
\label{sec:enforcement}

In this section we explore how an existing enforcement mechanism called {\em
  Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
the formal stack-safety property described in the previous
section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not..
We later validate this claim using property-based random testing.

\paragraph*{A Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\cite{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
instruction opcode itself, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
Existing work \cite{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
has shown that a wide range of
micro-policies can be defined using this scheme.

@@

\medskip
\paragraph*{A Lazy Alternative}

>>>>

Our threat model is very strong, allowing the attacker (i.e., the
callee) to execute arbitrary code, including attempting to smash
the stack to disrupt the program’s control flow. Our stack safety
properties demand that, even in the presence of such attackers, confidentiality
and integrity still apply to data in suspended stack frames
until the callee returns.
%
We can also, optionally, enforce {\em well-bracketed control
flow (WBCF)}~\cite{SkorstengaardSTKJFP}---a global requirement that
callees must always return to their immediate caller, if they return
at all.
\apt{Originally, we were using this spot to contrast with Skorstengaard et al. in the intro.
  Do we really need to do that? Reads better without!}

@@

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security, combine these with an existing notion of well-bracketed
control flow to formalize a {\em stepwise stack-safety}
property, and which captures the intuitive concept of
strict stack safety enforced over each step of execution~(\cref{sec:prelim,sec:lse}).
As a case study, we explore how our properties are enforced 
by the {\em Depth Isolation} micro-policy
of Roessler and DeHon\cite{DBLP:conf/sp/RoesslerD18}, but not by some of its optimized variants~(\cref{sec:enforcement}).
\item We formalize weaker but more extensional \emph{observational} variants of
  confidentiality and integrity. Both are parameterized over a notion of external
  observation, and are violated only if accessing secrets or overwriting
  data causes a visible change in the system behavior (\cref{sec:lazy}).
\item We show that our definitions can be
extended to describe argument passing on the stack, callee-saves registers, and a simple coroutine system (\cref{sec:ext}).
\item We use property-based random testing to validate the relationship between
  our properties and micro-policies; in the process, we extend the generation by
execution techniques of Hri\c{t}cu et al.\cite{TestingNI:ICFP} to handle nested
properties (\cref{sec:testing})\apt{Seems like some of the relevant support for this contribution is now cut?}
\end{itemize}
\Cref{sec:relwork,sec:future} discuss related and future work.

@@

Our model is flexible enough to support many realistic uses of the stack.
We intentionally say little or nothing about other structures
such as the heap.  Heap safety has already been studied, in isolation,
by Azevedo de Amorim et al.~\cite{DBLP:conf/post/AmorimHP18}, and we expect that their correctness condition and ours can be
combined cleanly to characterize correct protection of stack + heap.

We do not currently support:
1) address-taken locals, as a satisfactory treatment of these seems to require
integration with a more general model of memory safety,
2) exceptions, \sna{I could just write up how to do these...}
3) tail calls,
4) dynamic code generation, which is difficult to integrate with code annotations in general, or
5) concurrency --- we can handle a limited form of coroutines, but
  not full-blown preemptive concurrency.

@@

\medskip
\section{Case Study: Micro-Policies}
\label{sec:enforcement}

In this section we explore how an existing enforcement mechanism called {\em
  Depth Isolation}~\cite{DBLP:conf/sp/RoesslerD18} implements
the formal stack-safety property described in the previous
section, and why its optimized form, {\em Lazy Tagging and Clearing}, does not.
We validate this claim using property-based random testing in \cref{sec:testing}.

\paragraph*{Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\cite{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
instruction opcode itself, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
Existing work \cite{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
has shown that a wide range of
micro-policies can be defined using this scheme.

@@

\medskip
\paragraph*{Lazy Alternative}
